% general sample of a presentation

%%% DOCUMENT TYPE
\documentclass{beamer}

%%% PRESENTATION THEME
\usetheme{Madrid}
\setbeamercovered{invisible}
\setbeamertemplate{navigation symbols}{}

%%% PACKAGES
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage[sort,compress,comma,super]{natbib}
\bibliographystyle{apalike}

%%% CUSTOM COMMANDS
% for natbib & beamer to kill error
\def\newblock{\hskip .11em plus .33em minus .07em}

%%% TITLE INFO
\title{Sample Presentation}
\author{Sean McKenna}
\institute[University of Utah]{ University of Utah \\ \medskip {\emph{mckennapsean@gmail.com}} }
\date{\today}

%%% BEGIN PRESENTATION
\begin{document}


\begin{frame}
  \titlepage
\end{frame}

\begin{frame}
  \frametitle{Introduction}
  \begin{itemize}
    \item briefly explain and introduce topic
  \end{itemize}
\end{frame}

\begin{frame}
  \frametitle{Background}
  \begin{itemize}
    \item important information to understand the topic
  \end{itemize}
\end{frame}

\begin{frame}
  \frametitle{Research}
  \begin{block}{Fundamental Research Question}
    What were you trying to find out?
  \end{block}
  \begin{block}{Significance}
    So what? Why is this important?
  \end{block}
\end{frame}

\begin{frame}
  \frametitle{Experiment}
  \begin{itemize}
    \item what did you do to answer your research question?
  \end{itemize}
\end{frame}

\begin{frame}
  \frametitle{Results}
  \begin{itemize}
    \item what did you get from your experiment?
  \end{itemize}
\end{frame}

\begin{frame}
  \frametitle{Analysis}
  \begin{itemize}
    \item what did you find from the results of your experiment?
  \end{itemize}
\end{frame}

\begin{frame}
  \frametitle{Conclusion}
  \begin{itemize}
    \item so what? why are your results important?
  \end{itemize}
\end{frame}

\begin{frame}
  \frametitle{Future Work}
  \begin{itemize}
    \item what are the next steps for research?
  \end{itemize}
\end{frame}

\begin{frame}
  \frametitle{References}
  \bibliography{sources}
\end{frame}


\end{document}
